\documentclass[11pt]{article}
        % Use font 11pt Roman.
%\documentstyle{article}

\usepackage{url}

%\def\references{\newpage\section*{References}}
%\newcounter{refcount}
%\def\lit#1{~[\ref{#1}]}
%\def\cit#1#2#3#4{#1\quad#2\quad{\it #3}\quad #4\vskip.15in }
%\def\citb#1#2#3{#1\quad{\it #2}.\quad#3.}
%\def\citu#1#2{#1\quad#2.}


\newcommand{\comment}[1]{}


% No header or foot; top 2.0in, left 0.6in, right 0.9in, bottom 0.5in

%\pagestyle{empty}
%\topmargin=-.5in
%\textheight=8.8in
%\leftmargin=-1.0in
%\oddsidemargin=.25in
%\oddsidemargin=.25in
%\textwidth=6.2in

\setlength{\textheight}{22.8true cm}
%\setlength{\textwidth}{15.4true cm}
\setlength{\textwidth}{15.9true cm}
\setlength{\oddsidemargin}{.55true cm}
\setlength{\evensidemargin}{.55true cm}
\setlength{\topmargin}{-1.0cm}
\pagestyle{empty}

\newlength{\oldparindent}
\setlength{\oldparindent}{\parindent}
\setlength{\parindent}{0pt}

% No indentation at the beginning of paragraphs.

% Don't hyphenate.
%\hyphenpenalty=10000


% Rsection: a resume section environment.
\newenvironment{Rsection}[1]{
  \noindent{\bf #1}
  \begin{list}{}{\topsep=-\parskip
                 \leftmargin=-0.0in}
    \item[]
}{
  \end{list}
%  \vspace \baselineskip
}
                           
% Section: a resume section environment.
\newenvironment{Section}[1]{
  \centerline{\bf #1}
  \begin{list}{}{\topsep=-\parskip
                 \leftmargin=-0.5in}
    \item[]
}{
  \end{list}
%  \vspace \baselineskip
}

% edlist: a resume education list environment.
\newenvironment{edlist}{
  \begin{list}{}{\topsep=-\parskip
                 \itemsep=0.5\baselineskip
                 \leftmargin=0.25in}
}{
  \end{list}
}


% edlist2: a resume education list environment.
\newenvironment{edlist2}{
  \begin{list}{}{\topsep=-\parskip
                 \itemsep=0.5\baselineskip
                 \leftmargin=0.25in
                 \setlength{\rightmargin}{\leftmargin}
                 \textwidth=4in}
}{
  \end{list}
}
                           

% honorlist: a resume honor list environment.
\newenvironment{honor2list}{
  \begin{list}{}{\topsep=-\parskip
                 \itemindent=-0.in
                 \itemsep=0.0\baselineskip
                 \leftmargin=.25in}
}{
  \end{list}
}  

\newenvironment{honor2enum}{
  \begin{enumerate}{}{\topsep=-\parskip
                 \itemindent=-0.in
                 \itemsep=0.0\baselineskip
                 \leftmargin=.25in}
}{
  \end{enumerate}
}  

\newenvironment{honor3enum}{
  \begin{enumerate}{}{\topsep=-\parskip
                 \itemindent=-0in
                 \itemsep=0.0\baselineskip
                 \leftmargin=.25in}
}{
  \end{enumerate}
}  



% honorlist: a resume honor list environment.
\newenvironment{honorlist}{
  \begin{list}{}{\topsep=-\parskip
                 \itemindent=-0.in
                 \itemsep=0.2\baselineskip
                 \leftmargin=.25in}
}{
  \end{list}
}  


% joblist: a resume job list environment.
\newenvironment{joblist}{
  \begin{list}{}{\topsep=-\parskip
                 \itemsep=0.5\baselineskip
                 \leftmargin=0.5in}
}{
  \end{list}
}

                                      
% job: a job entry for a joblist environment
\newcommand{\job}[3]{
  \item[]
  \begin{tabbing}
    \hspace{-0.5in}\=\hspace{1.5in}\=\kill
    \>    {\it #1:}       \> #2   \\
    \>                    \> #3   \\
  \end{tabbing}
  \vspace{-\baselineskip}
}

              
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\leftmargin=-1in
\begin{document}


\centerline{\Large \bf {\LARGE \bf D}AVID {\LARGE \bf W}ALKER}
\vskip .3in

\normalsize
\it
\begin{tabular}{@{\hspace{0in}}l@{\hspace{1.4in}}l}
35 Olden Street   & Phone: (609) 258--7654  \\
Princeton University   & Email: {\rm dpw@cs.princeton.edu} \\
Princeton, NJ 08544

\end{tabular}
\rm
\vskip.3in

\begin{Rsection}{\Large \bf {Professional Preparation}}
\vskip.1in
\begin{edlist}

\item {\bf Queens University, Kingston, Ontario, Canada }
  \begin{honor2list}
  \item {Computer Science, B.Sc. Honors, 1995.}  

  \end{honor2list}
\item {\bf Cornell University}
\begin{honor2list}
\item {Computer Science, M.S. 1999.}
\item {Computer Science, Ph.D. 2001.}
\end{honor2list}
\end{edlist}
\end{Rsection}


\vskip.3in
\begin{Rsection}{\Large \bf {Appointments}}
\vskip.1in
\begin{edlist}
\item {\bf Princeton University}\\
{\bf  Department of Computer Science} \\
%\vskip .05in
\vspace{-1ex}
\begin{honor2list}
\item Assistant Professor, Feb. 2002--July 2008.
\item Associate Professor with Tenure, July 2008--Present.
\end{honor2list}

\item {\bf Carnegie Mellon University}\\
{\bf  Department of Computer Science} \\ 
%\vskip .05in
\vspace{-1ex}
\begin{honor2list}
\item Postdoctoral Fellow, Oct. 2000--Oct. 2001
\end{honor2list}


%\newpage
\end{edlist}
\end{Rsection}


\vskip.3in
\begin{Rsection}{{\Large \bf {Selected Publications}} 
}
\vskip.1in
\begin{edlist}

\item {\bf Selected papers of greatest relevance (in chronological order)}
\begin{honor2enum}
\vskip .1in

\item LearnPADS: Fully Automatic Tool Generation From Ad Hoc
Data. Kathleen Fisher, David Walker and Kenny Q. Zhu.  ACM SIGMOD Demo
Session.  June 2008.

\item From Dirt to Shovels: Fully Automatic Tool Generation From Ad
Hoc Data. Kathleen Fisher, David Walker, Kenny Q. Zhu and Peter White.
ACM SIGPLAN-SIGACT Symposium on Principles of Programming Languages.
January 2008. (POPL 08)

\item PADS/ML: A Functional Data Description Language.  Yitzhak
Mandelbaum, Kathleen Fisher, David Walker, Mary Fernandez, and Artem
Gleyzer.  ACM SIGPLAN-SIGACT Symposium on Principles of Programming
languages.  January 2007.  (POPL 07)

\item PADS: An End-to-end System for Processing Ad Hoc Data.  Mark Daly, Mary Fernandez, Kathleen Fisher, Robert Gruber, Yitzhak Mandelbaum, David Walker and Xuan Zheng.  ACM SIGMOD Demo Session.  June 2006. 

\item The Next 700 Data Description Languages. Kathleen Fisher, 
Yitzhak Mandelbaum and David Walker.  ACM SIGPLAN-SIGACT Symposium 
on Principles of Programming languages.  January 2006. (POPL 06)
%

\end{honor2enum}

\newpage
\item {\bf Selected other publications (in chronological order):}
\begin{honor2enum}

\item Composing Expressive Run-time Security Policies.  Lujo Bauer,
Jay Ligatti, and David Walker.  ACM Transactions on Software
Engineering and Methodology, Volume 18, Issue 3, pp 9:1-9:43, May
2009.

\item Run-time Enforcement of Nonsafety Policies. Jay Ligatti, Lujo Bauer, and David Walker. ACM Transactions on Information and System Security, Volume 12, Issue 3, pp 1-41, January 2009.

\item Fault-tolerant Typed Assembly Language. Frances Perry, Lester
Mackey, George A. Reis, Jay Ligatti, David I. August, and David
Walker.  ACM SIGPLAN Conference on Programming Language Design and
Implementation. June 2007.  Winner of the PLDI 2007 Best Paper Award.

\item Substructural Type Systems.  David Walker.
Chapter 1 of Benjamin Pierce, ed.,
Advanced Topics in Type Systems for Programming Languages.  
MIT Press, January 2005.

\item
From System F to Typed Assembly Language.  Greg Morrisett, David
Walker, Karl Crary, and Neal Glew. Twenty-Fifth ACM SIGPLAN Symposium
on Principles of Programming Languages. pages 85-97, San Diego,
January 1998.  Winner of the Most Influential 1998 POPL Paper Award
(awarded after 10 years reflection at POPL 2008).

\end{honor2enum}

\end{edlist}
\end{Rsection}



\vskip.3in
\begin{Rsection}{\Large \bf {Synergistic Activities}}
\begin{honor2list}
\vskip .1in
\begin{itemize}
\item Core member of the Computing Community Consurtium (CCC) 
Visioning Study on cross-layer reliability: System-level, Cross-layer 
Cooperation to Achieve Predictable Systems from Unpredictable Components.
\item Summer School on Programming Languages and Security.  
Steering Committee.
\item Program committee member for ACM SIGPLAN International Conference on
Programming Language Design and Implementation, 2010. (PLDI 2010),   
ACM SIGPLAN International Conference on Object-Oriented Programming,
Systems,  Languages, and Applications, 2009. (OOPSLA 2009) and the
21st Symposium on Implementation and Application of Functional 
Languages (IFL 2009).
%\item 
%Program Chair for FOAL 2005, SPACE 2006.  Program committee
%for ICFP 2004, TLDI 2005, FOAL 2006, ICALP 2006, POPL 2007.  
%Periodic program chair and host for
%the New Jersey Programming Languages Seminar, 2002-now.
%\item Publicity chair for ACM SIGPLAN-SIGACT Symposium on
%Principles of Programming Languages, 2003--2006.
\end{itemize}



\end{honor2list}
\end{Rsection}
\vskip.3in
\begin{Rsection}{\Large \bf {Collaborators and Other Affiliations}}
\begin{itemize}
\vskip .1in
\item Non-Princeton Collaborators: 
Amal Ahmed (Indiana U),
%Andrew Appel (Princeton),
Zena Ariola (U of Oregon),
Lujo Bauer (Carnegie Mellon),
%Iliano Cervesato (Tulane),
Luis Ceze (U of Washington),
%Karl Crary (Carnegie Mellon University),
%Daniel Dantas (Princeton University),
%Ed Felten (Princeton University),
Mary Fernandez (AT\&T),
Kathleen Fisher (AT\&T),
Neal Glew (Intel),
Daniel Grossman (U of Washington),
Robert Harper (Carnegie Mellon),
Limin Jia (Carnegie Mellon),
Trevor Jim (AT\&T),
Jarred Ligatti (U of South Florida),
Boon Thau Loo (U of Pennsylvania),
Yitzhak Mandelbaum (AT\&T),
Greg Morrisett (Harvard),
Benjamin Pierce (U of Pennsylvania),
Frank Pfenning (Carnegie Mellon),
Jonathan Smith (U of Pennsylvania),
%Fred Smith (MathWorks),
%Frances Spalding (Princeton),
%Daniel Wang (Princeton),
%Geoffrey Washburn (U of Pennsylvania),
%Kevin Watkins (Carnegie Mellon),
Stephanie Weirich (U of Pennsylvania),
Matt Welsh (Harvard),
Steve Zdancewic (U of Pennsylvania)

\item Graduate (Ph.D.) Advisor: 
Greg Morrisett (Cornell University, now Harvard University)
\item Thesis Advisees (graduated): Dan Dantas, Limin Jia, Jarred Ligatti, 
Yitzhak Mandelbaum, Frances Spalding 
\item Thesis Advisees (current):  Christian Bell, Matthew Meola, Qian Xi
\item Postdoctoral Advisees (current):  Kenny Q. Zhu
\end{itemize}
\end{Rsection}



% \iffalse
% \vskip.3in
% \begin{Rsection}{\Large \bf {Selected Awards and Fellowships}}
% \begin{itemize}
% \vskip .1in
% \end{itemize}
% \end{Rsection}
% \fi




%\newpage

%\newpage





%%%%%$$^{\ref{fn:second}}$







%\vskip.3in




\end{document}

