\section{Conclusion} 
\label{sec:conclusion}

Ad hoc data is pervasive and valuable: in industry, in medicine, and
in scientific research.  Such data tends to have poor documentation,
to contain various kinds of errors, and to be voluminous.  Unlike
well-behaved data in standardized relational or \xml{} formats, such
data has little or no tool support, forcing data analysts and
scientists to waste valuable time writing brittle custom code, even if
all they want to do is convert their data into a well-behaved format.
To improve the situation, various researchers have developed data
description languages such as \pads{}, \datascript{}, and
\packettypes{}.  Such languages allow analysts to write terse,
declarative descriptions of ad hoc data.  A compiler then generates a
parser and customized tools.  Because these languages are tailored to
their domain, they can provide useful services automatically while a
more general purpose tool, such as \lex{}/\yacc{} or \perl{}, cannot.

In the spirit of Landin, we have taken the first steps toward
specifying a semantics for this class of languages by defining the
data description calculus \ddc{}.  This calculus, which is a dependent
type theory with a simple set of orthogonal primitives, is expressive
enough to describe the features of \pads{}, \datascript{}, and
\packettypes{}.  In keeping with the spirit of the data description
languages, our semantics is transformational: instead of simply
recognizing a collection of input strings, we specify how to transform
those strings into canonical in-memory representations annotated with
error information.  Furthermore, we prove that the error information
is meaningful, allowing analysts to rely on the error summaries rather
than having to re-vet the data by-hand.

We have already used the semantics to identify bugs in the
implementation of \padsc{} and to highlight areas where \padsc{}
sacrifices safety for speed.  We have also used the semantics as a guide
for the design of a whole new language, \padsml{}, designed
specifically for functional programmers.  In the future, we hope 
\ddc{} will serve as a solid foundation for the next 700 data 
description languages to come.
