%\documentclass[fleqn]{article}
\documentclass[tfpsymp,ams,pagenumbers]{tfp07symp}
\setcounter{paper}{13} 
\usepackage{xspace,pads,math-cmds,
            math-envs,inference-rules,times,
            verbatim,alltt,multicol,proof}

\usepackage{code} 
\usepackage{epsfig}
% \usepackage{natbib}

% to do:
% 1. related work, including Eelco Visser's work and John Hughes
% pretty printers. First, mention that printing and pretty printing
% are different problems. Then, differentiate semantics and implementation.
% 2. add abstract
% 3. mention round tripping in future work.
% 4. Include full figures.

%\setlength{\oddsidemargin}{0in}
%\setlength{\evensidemargin}{0in}
%\setlength{\textwidth}{6.5in}
%\setlength{\textheight}{8.5in}

\begin{document}
\title{A Dual Semantics for the Data Description Calculus\\(Extended Abstract)}
\author{Kathleen Fisher$^1$ \qquad Yitzhak Mandelbaum$^{1,2}$ \qquad David Walker$^2$}
\institute{AT\&T Research \and Princeton University}

\input{definitions}

\maketitle{}

\begin{abstract}  
  In previous work, we presented \ddc{}, a semantic framework for
  understanding {\em Data Description Languages}, a class of
  domain-specific languages for declaratively describing data formats
  for the purpose of automatically constructing format-specific
  data-processing tools.  However, our initial work on \ddc{} told
  only a fraction of the semantic story concerning data description
  languages.  Many data description languages not only provide
  parsers, but also other tools.  Amongst the most common auxiliary
  tools are printers, as reliable communication between programs
  depends upon both input (parsing) and output (printing).  In this
  work, we have defined the semantics of printers for \ddc{}, thereby
  specifying more completely the relationship between raw data and
  in-memory data, for any given format described in \ddc{}.  We also
  prove a collection of theorems for the new semantics that serve as
  duals to our theorems concerning parsing.  This new printing
  semantics has many of the same practical benefits as our older
  parsing semantics: We can use it as a check against the correctness
  of our printer implementations and as a guide for the implementation
  of future data description languages.
\end{abstract}

\input{intro}

\input{ddcpoly}

\input{related_work}

\input{conc}

\paragraph*{Acknowledgments:} This material is based upon work supported by the National Science
   Foundation under Grant Nos. 0615062 and 0612147.
Any opinions, findings, and conclusions or recommendations
   expressed in this material are those of the authors and do not
   necessarily reflect the views of the NSF.

% \bibliographystyle{abbrv}
% \bibliographystyle{plainnat}
\bibliographystyle{alpha}

  \bibliography{pads} 
%\bibliography{../common/pads-long,../common/pads}

\end{document}

%%% Local Variables:
%%% mode: outline-minor
%%% End:

