%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \centerline{
% \begin{tabular}{cc}
%           \\[-1ex]
% \multicolumn{2}{c}{\Large{} Tools for Processing Ad Hoc Data Sources} \\
%           & \\[-1ex]
% Kathleen Fisher & David Walker (PI)\\
% AT\&T  & Princeton University\\[2ex]
% \end{tabular}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\paragraph*{Intellectual Merits:} 
An {\em ad hoc data format} is any nonstandard data format for which
parsing, querying, analysis or transformation tools are not readily
available.  \xml{} is not an ad hoc data format --- there are hundreds
of tools for reading, querying, and transforming \xml{}.   However,
network administrators, computer scientists, information systems managers,
financial analysts, biologists, chemists, and physicists deal with ad hoc
data in a myriad of complex formats on a daily basis.
This data, which is often unpredictable, poorly documented,
filled with errors (malicious and benign) and unwieldy,
poses tremendous challenges to its users and the software
that manipulates it.  

Our end goal is to help software developers 
create efficient, reliable and easy-to-maintain software for
for processing ad hoc data.  We will achieve our goal
through the design and implementation of \datatype,
a revolutionary new language with intrinsic support for 
processing ad hoc data.  \datatype{} will have 
two logical components.

\begin{enumerate}
\item A Universal Data Description Language.  We will aim to develop a 
high-level, easy-to-understand description language for \datatype, capable of
concisely and accurately describing the physical format of
any ad hoc data source. 
\item An Expressive and Efficient Ad Hoc Data Transformation Language.
The data transformation language will allow programmers
automatically generating a suite of high-assurance,
efficient data processing tools that make understanding, manipulating 
and transforming ad hoc data an effortless task.
\end{enumerate}

% An {\em ad hoc data format} is any nonstandard data format
% for which parsing, querying, analysis or transformation
% tools are not readily available.
% There are vast amounts of useful data 
% stored in traditional databases and \xml\ formats, but there is just
% as much in ad hoc formats.
% In the networking and telecommunications
% domain, we find ad hoc formats for web server logs, 
% netflows capturing internet traffic, 
% log files characterizing IP backbone resource utilization,
% wire formats for legacy telecommunication billing systems, 
% and telephone call detail data, to name just a few.
% The ability to parse, query, monitor and transform this ad hoc data
% is a crucial part of maintaining and providing security for
% our communications infrastructure.

% Explanation as to why ad hoc data is relevant to cyber trust goes here.

% Processing ad hoc data is challenging for a variety of
% reasons. First, ad hoc data arrives ``as is:'' an analyst might want the
% data delivered in \xml, but unfortunately, he or she has just got to deal
% with it as it is.
% Second, documentation for the format may not exist at all, or it may be
% out of date.  
% Third, ad hoc data frequently contain errors, for a variety of
% reasons: malfunctioning equipment, race conditions on log
% entry~\cite{wpp}, malicious agents attempting to exploit
% software vulnerabilities, human error in entering data, and others. 
% The appropriate response to such errors depends on the application. 
% % Some applications require the data to be error free: 
% % if an error is detected, processing needs to stop immediately and a human
% % must be alerted.  Other applications can repair the data, while still
% % others can simply discard erroneous or unexpected values.  
% For some applications, errors or anomalous data can be the most 
% interesting part because they can signal potential network
% vulnerabilities, opportunities or ongoing fraud in telephony systems, 
% or failure to communicate between systems.  A fourth challenge is that 
% ad hoc data sources can be high volume.  For instance,
% AT\&T's call-detail stream contains roughly 300~million calls per day
% requiring approximately 7GBs of storage space.

Work has begun on developing a high-level specification 
language called \pads{} for describing ad hoc data formats.  From a
\pads{} specification it is currently possible to generate a number of
tools for processing that data automatically including (1)
efficient parsers for reading ad hoc data, (2) an interface to a 
general-purpose data query engine, (3) an analyzer that generates 
statistical summaries of data read by the parser, and (4)
a translator from the ad hoc format into \xml.  

While the work to date demonstrates the feasibility of the \pads{}
approach, the \pads{} design and implementation are still in their infancy:
We have discovered many common, real-world data formats that the current PADS 
infrastructure is incapable of describing, parsing or analyzing. 
To address these deficiencies, we propose to explore innovative ways
of extending the basic \pads{} specification language and related
infrastructure.  The second thrust of our research involves
improving the automatic tool generation infrastructure in the \pads{} system.
The current tool generation system is brittle, inflexible and suffers from
subpar performance.  We will develop a novel extension to \pads{} 
that allows users to augment \pads{} descriptions with application-specific
attributes and customizations.  The attributes convey semantic information to
the tool generators and the customizations improve error-handling
and performance.  Our new architecture has the promise of being
substantially more robust, flexible and efficient.
Finally, in order to improve the reliability of the infrastructure,
we will formalize \pads{} and prove important correctness properties
of our formal model.  Overall, this proposal involves challenging research in
language design, efficient systems implementation and
semantic analysis, all aimed at solving real-world data processing 
problems.

\paragraph*{Broader Impacts:}  We are collaborating with researchers at
AT\&T who will be able to use our tools to address real problems such as
telephone fraud detection.  In addition, our tools will be 
freely available to researchers and scientists over the web.  
Moreover, part of our mission will be to work with 
biologists, chemists and physicists at Princeton and the broader community
to help them with their data 
processing needs.  We have already been meeting with Olga Troyanskaya,
who works in Princeton's Lewis-Sigler Institute for Integrative Genomics 
on pathway modeling and analysis of protein-protein interactions.  Many of our proposed
extensions to \pads{} were inspired directly by her data processing needs.
The collaboration between Computer Science and Genomics will also be an excellent platform
for developing interdisciplinary undergraduate research projects.
