%\documentclass[fleqn]{article}
%\documentclass[nocopyrightspace]{sigplanconf}
\documentclass{sigplanconf}

\usepackage{xspace,pads,amsmath,math-cmds,
            math-envs,inference-rules,times,
            verbatim,alltt,multicol,proof,url}

\usepackage{code} 
\usepackage{epsfig}

%\setlength{\oddsidemargin}{0in}
%\setlength{\evensidemargin}{0in}
%\setlength{\textwidth}{6.5in}
%\setlength{\textheight}{8.5in}

\begin{document}
\conferenceinfo{POPL'07} {January 17--19, 2007, Nice, France.}

\copyrightyear{2007}

\copyrightdata{1-59593-575-4/07/0001} 

\title{PADS/ML: A Functional Data Description Language}
\authorinfo{Yitzhak Mandelbaum$^\ast$, Kathleen Fisher$^\dagger$, David Walker$^\ast$, Mary
  Fernandez$^\dagger$, Artem Gleyzer$^\ast$}
{$^\ast$Princeton University \ \ \ \ \ \ \ $^\dagger$AT\&T Labs Research}
 {\mono{yitzhakm,dpw,agleyzer@CS.Princeton.EDU} \ \ \ \ \mono{kfisher,mff@research.att.com}}

\input{definitions}

\maketitle{}

\begin{abstract}  

  Massive amounts of useful data are stored and processed in
  \textit{ad hoc} formats for which common tools like parsers,
  printers, query engines and format converters are not readily 
  available.  In this paper,
  we explain the design and implementation of \padsml{},
  a new language and system that facilitates the
  generation of data processing tools for ad hoc formats.
  The \padsml{} design includes features such as
  dependent, polymorphic and recursive datatypes, which allow
  programmers to describe
  the syntax and semantics of ad hoc data in a concise,
  easy-to-read notation.
  The \padsml{} implementation compiles these descriptions
  into \ml{} structures and functors
  that include types for parsed data, functions for parsing
  and printing, and auxiliary support
  for user-specified, format-dependent and format-independent 
  tool generation.
%%  \edcom {sentence here about \padsml{} being based on theory?}
%% dpw -- couldn't think of anything useful to say.

\end{abstract}

\category{D.3.2}{Language Classifications}{Applicative (functional) languages}

\terms
Languages

\keywords
Data description languages, domain-specific languages, functional programming,
dependent types, ML, modules, parsing, printing


\input{intro}

\input{pml_overview}

\input{impl}

\input{gentool}

%\input{related_work}

% \input{conc}

\input{conc2}

\paragraph*{Acknowledgments.}
We would like to thank Derek Dreyer for discussions and advice on
advanced module systems.  Some of this material is based upon work 
supported by NSF awards 0615062 and 0612147.  
Any opinions, findings, and conclusions 
or recommendations expressed in the material are those of the authors 
and do not necessarily reflect the views of the NSF.

\bibliographystyle{abbrv}
\bibliography{pads}
%\bibliography{../common/pads-long,../common/pads}

%\input{appendix}

\end{document}

%%% Local Variables:
%%% mode: outline-minor
%%% End:

