\documentclass[11pt]{article}
        % Use font 11pt Roman.
%\documentstyle{article}

%\def\references{\newpage\section*{References}}
%\newcounter{refcount}
%\def\lit#1{~[\ref{#1}]}
%\def\cit#1#2#3#4{#1\quad#2\quad{\it #3}\quad #4\vskip.15in }
%\def\citb#1#2#3{#1\quad{\it #2}.\quad#3.}
%\def\citu#1#2{#1\quad#2.}


\newcommand{\comment}[1]{}


% No header or foot; top 2.0in, left 0.6in, right 0.9in, bottom 0.5in

%\pagestyle{empty}
%\topmargin=-.5in
%\textheight=8.8in
%\leftmargin=-1.0in
%\oddsidemargin=.25in
%\oddsidemargin=.25in
%\textwidth=6.2in

\setlength{\textheight}{22.8true cm}
%\setlength{\textwidth}{15.4true cm}
\setlength{\textwidth}{15.9true cm}
\setlength{\oddsidemargin}{.55true cm}
\setlength{\evensidemargin}{.55true cm}
\setlength{\topmargin}{-1.0cm}
\pagestyle{empty}

\newlength{\oldparindent}
\setlength{\oldparindent}{\parindent}
\setlength{\parindent}{0pt}

% No indentation at the beginning of paragraphs.

% Don't hyphenate.
%\hyphenpenalty=10000


% Rsection: a resume section environment.
\newenvironment{Rsection}[1]{
  \noindent{\bf #1}
  \begin{list}{}{\topsep=-\parskip
                 \leftmargin=-0.0in}
    \item[]
}{
  \end{list}
%  \vspace \baselineskip
}
                           
% Section: a resume section environment.
\newenvironment{Section}[1]{
  \centerline{\bf #1}
  \begin{list}{}{\topsep=-\parskip
                 \leftmargin=-0.5in}
    \item[]
}{
  \end{list}
%  \vspace \baselineskip
}

% edlist: a resume education list environment.
\newenvironment{edlist}{
  \begin{list}{}{\topsep=-\parskip
                 \itemsep=0.5\baselineskip
                 \leftmargin=0.25in}
}{
  \end{list}
}


% edlist2: a resume education list environment.
\newenvironment{edlist2}{
  \begin{list}{}{\topsep=-\parskip
                 \itemsep=0.5\baselineskip
                 \leftmargin=0.25in
                 \setlength{\rightmargin}{\leftmargin}
                 \textwidth=4in}
}{
  \end{list}
}
                           

% honorlist: a resume honor list environment.
\newenvironment{honor2list}{
  \begin{list}{}{\topsep=-\parskip
                 \itemindent=-0.in
                 \itemsep=0.0\baselineskip
                 \leftmargin=.25in}
}{
  \end{list}
}  

\newenvironment{honor2enum}{
  \begin{enumerate}{}{\topsep=-\parskip
                 \itemindent=-0.in
                 \itemsep=0.0\baselineskip
                 \leftmargin=.25in}
}{
  \end{enumerate}
}  

\newenvironment{honor3enum}{
  \begin{enumerate}{}{\topsep=-\parskip
                 \itemindent=-0in
                 \itemsep=0.0\baselineskip
                 \leftmargin=.25in}
}{
  \end{enumerate}
}  



% honorlist: a resume honor list environment.
\newenvironment{honorlist}{
  \begin{list}{}{\topsep=-\parskip
                 \itemindent=-0.in
                 \itemsep=0.2\baselineskip
                 \leftmargin=.25in}
}{
  \end{list}
}  


% joblist: a resume job list environment.
\newenvironment{joblist}{
  \begin{list}{}{\topsep=-\parskip
                 \itemsep=0.5\baselineskip
                 \leftmargin=0.5in}
}{
  \end{list}
}

                                      
% job: a job entry for a joblist environment
\newcommand{\job}[3]{
  \item[]
  \begin{tabbing}
    \hspace{-0.5in}\=\hspace{1.5in}\=\kill
    \>    {\it #1:}       \> #2   \\
    \>                    \> #3   \\
  \end{tabbing}
  \vspace{-\baselineskip}
}

              
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\leftmargin=-1in
\begin{document}


\centerline{\Large \bf {\LARGE \bf D}AVID {\LARGE \bf W}ALKER}
\vskip .3in

\normalsize
\it
\begin{tabular}{@{\hspace{0in}}l@{\hspace{1.4in}}l}
35 Olden Street   & Phone: (609) 258--7654  \\
Princeton University   & Email: {\rm dpw@cs.princeton.edu} \\
Princeton, NJ 08544

\end{tabular}
\rm
\vskip.3in

\begin{Rsection}{\Large \bf {Professional Preparation}}
\vskip.1in
\begin{edlist}

\item {\bf Queens University, Kingston, Ontario, Canada }
  \begin{honor2list}
  \item {Computer Science, B.Sc. Honors, 1995.}  

  \end{honor2list}
\item {\bf Cornell University}
\begin{honor2list}
\item {Computer Science, M.S. 1999.}
\item {Computer Science, Ph.D. 2001.}
\end{honor2list}
\end{edlist}
\end{Rsection}


\vskip.3in
\begin{Rsection}{\Large \bf {Appointments}}
\vskip.1in
\begin{edlist}
\item {\bf Princeton University}\\
{\bf  Department of Computer Science} \\
%\vskip .05in
\vspace{-1ex}
\begin{honor2list}
\item Assistant Professor, Feb. 2002--present.
\end{honor2list}

\item {\bf Carnegie Mellon University}\\
{\bf  Department of Computer Science} \\ 
%\vskip .05in
\vspace{-1ex}
\begin{honor2list}
\item Postdoctoral Fellow, Oct. 2000--Oct. 2001
\end{honor2list}


%\newpage
\end{edlist}
\end{Rsection}


\vskip.3in
\begin{Rsection}{{\Large \bf {Selected Publications}} 
}
\vskip.1in
\begin{edlist}

\item {\bf Selected papers of greatest relevance (in chronological order)}
\begin{honor2enum}
\vskip .1in

\item Harmless advice.  Daniel S. Dantas and David Walker.
To appear in Workshop on
Foundations of Object-oriented
programming.  January 2005.

\item Edit Automata: Enforcement Mechanisms for Run-time Security Policies.  Jay Ligatti, Lujo Bauer and David Walker.  
International Journal of Information Security. 2004

\item A Theory of Aspects.  David Walker, Steve Zdancewic and Jay Ligatti.  In the ACM SIGPLAN International Conference on Functional Programming, August 2003. 

\item Types and Effects for Non-interfering Program Monitors.  
Lujo Bauer, Jarred Ligatti and David Walker.  
International Symposium on Software Security.  Tokyo, November, 2002.  

\item A Type System for Expressive Security Policies.  David Walker. 
Twenty-Seventh ACM SIGPLAN Symposium on Principles of Programming Languages. 
pages 254-267, Boston, January 2000

%


\end{honor2enum}
\item {\bf Selected other publications (in chronological order):}
\begin{honor2enum}

\item Substructural type systems.  David Walker.
Chapter 1 of Benjamin Pierce, ed.,
Advanced Topics in Type Systems for Programming Language.  
To appear January 2005.

\item An Effective Theory of Type Refinements.  Yitzhak Mandelbaum, David Walker and Robert Harper.  In the ACM SIGPLAN International Conference on Functional Programming, August 2003. 

\item 
Greg Morrisett, Karl Crary, Neal Glew, and David Walker. Stack-based
Typed Assembly Language.  Journal on Functional Programming,
12(1):43-88, January 2002.

\item 
David Walker, Karl Crary, and Greg Morrisett. Typed Memory Management
via Static Capabilities. ACM Transactions on Programming Languages and
Systems, 22(4):701-771, July 2000.   A previous version of this paper
appeared in the 26th ACM Symposium on Principles of Programming Languages.

\item 
Greg Morrisett, David Walker, Karl Crary, and Neal Glew. From System F
to Typed Assembly Language. ACM Transactions on Programming Languages
and Systems, 21(3):528-569, May 1999.  A previous version of this paper
appeared in the 25th ACM Symposium on Principles of Programming Languages.


\end{honor2enum}

\end{edlist}
\end{Rsection}



\vskip.3in
\begin{Rsection}{\Large \bf {Synergistic Activities}}
\begin{honor2list}
\vskip .1in
\item Program Chair, Workshop on Foundations of Aspect-Oriented
programming languages.  March 2005.
\item Alfred P. Sloan Fellow.  Sept 2004-Sept 2006. 
\item Program Committee,  ACM SIGPLAN International Conference
on Functional Programming.  September 2004.
\item Program Committee,  Workshop on Foundations of Aspect-Oriented
programming languages.  March 2004.
\item ACM SIGPLAN Summer School on Software Security.  Chair.  June 2004.
\item 
CAREER:  Programming Languages for Secure and Reliable Component Software Systems.  NSF Career Grant, CCR-0238328,  PI.  July 2003-June 2008.  
\item Publicity chair and executive committee for ACM Symposium on
Principles of Programming Languages, 2003--2006.




\end{honor2list}
\end{Rsection}
\vskip.3in
\begin{Rsection}{\Large \bf {Collaborators and Other Affiliations}}
\begin{itemize}
\vskip .1in
\item Collaborators: 
Amal Ahmed (Harvard University),
%Andrew Appel (Princeton University),
Zena Ariola (University of Oregon),
Lujo Bauer (Carnegie Mellon University),
Iliano Cervesato,
Karl Crary (Carnegie Mellon University),
%Daniel Dantas (Princeton University),
%Ed Felten (Princeton University),
Kathleen Fisher (AT\&T),
Neal Glew (Intel),
Daniel Grossman (University of Washington),
Robert Harper (Carnegie Mellon University),
%Limin Jia (Princeton University),
%Jarred Ligatti (Princeton University),
%Yitzhak Mandelbaum (Princeton University),
Greg Morrisett (Harvard University),
Benjamin Pierce (University of Pennsylvania),
Frank Pfenning (Carnegie Mellon University),
Fred Smith (MathWorks),
%Frances Spalding (Princeton University),
%Daniel Wang (Princeton University),
Geoff Washburn (University of Pennsylvania),
Kevin Watkins (Carnegie Mellon University),
Stephanie Weirich (University of Pennsylvania),
Steve Zdancewic (University of Pennsylvania)

\item Graduate (Ph.D.) Advisor: 
Greg Morrisett (Cornell University, now Harvard University)
\item Thesis Advisees (graduated): none
\item Thesis Advisees (current): Dan Dantas, Limin Jia, Jarred Ligatti, 
Yitzhak Mandelbaum, Frances Spalding
\end{itemize}
\end{Rsection}



% \iffalse
% \vskip.3in
% \begin{Rsection}{\Large \bf {Selected Awards and Fellowships}}
% \begin{itemize}
% \vskip .1in
% \end{itemize}
% \end{Rsection}
% \fi




%\newpage

%\newpage





%%%%%$$^{\ref{fn:second}}$







%\vskip.3in




\end{document}

