\section{Conclusions}
\label{sec:future}

\padsml{} is a high-level, domain-specific language and system
designed to help improve the productivity of the legions of data analysts
who work with ad hoc data on a regular basis.  
Inspired by the type structure of functional
programming languages, \padsml{} uses dependent,
polymorphic and recursive data types to describe the syntax and 
the semantic properties of ad hoc data sources.  The language 
is compact and expressive, capable of describing data from diverse 
domains including networking, computational biology, finance, and physics. 
The \padsml{} compiler uses a ``types as modules'' compilation strategy
in which every \padsml{} type definition is compiled into
an \ocaml{} module containing types for data representations
and functions for data processing.  Functional programmers
can use the generated modules to write clear and concise {\em format-dependent}
data processing programs.  Furthermore, our system design
allows external tool developers to write new {\em format-independent} tools
simply by supplying a module that matches the appropriate generic
signature.  The latest release of our implementation is available
at \url{http://www.padsproj.org/padsml/}.

The next step in our long-term agenda is to build a new generation of
format-independent data analysis tools.  While our current tools
perform some simple syntactic analysis and transformation, we intend
our next generation toolkit to perform deeper semantic analysis and
more sophisticated transformations.  For example, we may explore
specification-driven, content-based search, clustering, test data
generation, machine learning, security, and data visualization.  We
believe that if we can automatically generate stand-alone, end-to-end
tools that perform these functions over arbitrary data, we can have a
substantial impact on the productivity of researchers in a broad array
of scientific fields ranging from computational biology through
computer science to cosmology and beyond.
