%\documentclass{acm_proc_article-sp-sigmod06}
\documentclass{sig-alternate-sigmod06}
\usepackage{url}
\usepackage{stmaryrd}
\usepackage{epsfig}
\usepackage{alltt}
\usepackage{times}
\usepackage{code}
\usepackage{xspace}

% \pdfpagewidth=8.5in
% \pdfpageheight=11in

\renewcommand{\floatpagefraction}{0.9}
\renewcommand{\dbltopfraction}{0.9}
\renewcommand{\dblfloatpagefraction}{0.9}
\addtolength{\textfloatsep}{-8pt}
\addtolength{\dbltextfloatsep}{-8pt}

\include{defs}
\newcommand{\mono}[1]{\texttt{#1}}
\newcommand{\dibbler}{Sirius}
\newcommand{\ningaui}{Altair}
\newcommand{\darkstar}{Regulus}

\newcommand{\abstractdm}{abstract data model}
\newcommand{\concretedm}{concrete data model}
\newcommand{\typeddm}{type-specialized concrete data model}

\title{PADS: An End-to-end System for Processing Ad Hoc Data}

\numberofauthors{3} 
\author{\alignauthor Mark Daly\\
\affaddr{Princeton University}\\
\email{mdaly@princeton.edu}
\alignauthor Yitzhak Mandelbaum and David Walker\\
\affaddr{Princeton University}\\
\email{yitzhakm@cs.princeton.edu\\dpw@cs.princeton.edu}
\alignauthor Mary Fern\'andez and Kathleen Fisher\\
\affaddr{AT\&T Labs Research} \\
\email{mff@research.att.com\\kfisher@research.att.com}}

\eat{
\additionalauthors{Robert Gruber (Google, 
  {\texttt{gruber@google.com}}), while at AT\&T Labs and Xuan Zheng (Univ. of Michigan, 
  {\texttt{xuanzh@eecs.umich.edu}}), supported by 
	 AT\&T Labs and NSF DMS 0354600.}}

\date{\today}

\begin{document}

\maketitle
\eat{
\begin{abstract}
Enormous amounts of data exist in ``well-behaved'' formats such as
relational tables and XML, which come equipped with extensive tool
support.  However, vast amounts of data also exist in non-standard or
\textit{ad hoc} data formats, which often lack standard or extensible
tools. This deficiency forces data analysts to implement
their own tools for parsing, querying, and analyzing their ad hoc
data.  The resulting tools typically interleave parsing, querying, and
analysis, obscuring the semantics of the data format and making it
nearly impossible for others to resuse the tools.

This proposal describes \pads{}, an end-to-end system for processing
ad hoc data sources.  The core of \pads{} is a declarative
language for describing ad hoc data sources and a data-description
compiler that produces customizable libraries for parsing the ad hoc
data.  A suite of tools built around this core include statistical
data-profiling tools, a query engine that permits viewing ad hoc
sources as XML and for querying them with XQuery, and an interactive
front-end that helps users produce \pads{} descriptions quickly.

Details about the \pads{} system are reported in technical
papers~\cite{fernandez+:padx,fisher+:pldi05,fisher+:popl06}.  A
shorter version of this proposal was accepted to the PLAN-X 2006
workshop~\cite{daly+:launchpads}.  An open-source implementation of
\pads{} is available for download~\cite{padsmanual}.
\end{abstract}}

\input{intro}
\input{pads}

\noindent
\textbf{Additional Authors.}
Robert Gruber ({\small{\texttt{gruber@google.com}}}), Google,
while at AT\&T Labs.
Xuan Zheng ({\small{\texttt{xuanzh@eecs.umich.edu}}}), Univ. of Michigan, 
supported by AT\&T Labs and NSF DMS 0354600.

\bibliographystyle{abbrv}
\bibliography{../pads,../galax}

\end{document}

