\documentclass[natbib]{sigplanconf}
\usepackage{xspace,pads,amsmath,math-cmds,
            math-envs,inference-rules,times,
            verbatim,alltt,multicol,proof,url}
\usepackage{epsfig}
\usepackage{code} 

\begin{document}

\conferenceinfo{POPL'08,}{January 7-12,2008, San Francisco, California, USA.} 
\copyrightyear{2008} 
\copyrightdata{978-1-59593-689-9/08/0001} 

\title{From Dirt to Shovels}
\subtitle{Fully Automatic Tool Generation from Ad Hoc Data}

\authorinfo{Kathleen Fisher}{
	   AT\&T Labs Research}
       {\mono{kfisher@research.att.com}}
\authorinfo{David Walker \qquad Kenny Q. Zhu}{
	   Princeton University}
       {\mono{{dpw,kzhu}@CS.Princeton.EDU}}
\authorinfo{Peter White}{
	   Galois, Inc.}
       {\mono{peter@galois.com}}
\input{definitions}

\maketitle{}

\begin{abstract}  
An {\em ad hoc data source} is any semistructured data source for
which useful data analysis and transformation tools are not 
readily available.  Such data must be queried, transformed and displayed by
systems administrators, computational biologists, financial analysts
and hosts of others on a regular basis.  
In this paper, we demonstrate that it is possible to generate a suite
of useful data processing tools, including a semi-structured query
engine, several format converters, a statistical analyzer and data
visualization routines directly from the ad hoc data itself, 
without any human intervention.  
The key technical contribution of the work is a multi-phase algorithm
that automatically infers the structure of an ad hoc data source and
produces a format specification in the \pads{} data description
language.  
Programmers wishing to implement custom data analysis tools
can use such descriptions to generate printing and parsing libraries
for the data.  Alternatively,  our software infrastructure will
push these descriptions through the \pads{} compiler, creating
format-dependent modules that, when linked with format-independent
algorithms for analysis and transformation, result in
fully functional tools.  We evaluate the performance of
our inference algorithm, showing it scales linearly
in the size of the training data --- completing in seconds, as opposed
to the hours or days it takes to write a description by hand.
We also evaluate the correctness of the algorithm, demonstrating that 
generating accurate descriptions often requires less than 5\% of the
available data.
\end{abstract}

\category{D.3.m}{Programming languages}{Miscellaneous}

\terms
Languages, Algorithms

\keywords
Data description languages, grammar induction, tool generation, ad hoc data


\section {Introduction}
\label{sec:intro}
\input{intro}

\section{The Internal Format Description Language}
\label{sec:review}
\input{review}

\section{The Inference Algorithm}
\label{sec:inference}
\input{algorithm}

\section{Experimental Evaluation}
\label{sec:exp}
\input{eval}

\section{Discussion}
\label{sec:discussion}
\input{discussion}

\section{Related Work}
\label{sec:related}
\input{related}


\section{Conclusions}
\label{sec:conclusion}
\input{conclusion}

\section*{Acknowledgments}

Our work benefited greatly from thoughts and comments from
Alex Aiken, David Blei, David Burke, Vikas Kedia, John Launchbury, Chris Ramming, 
Rob Schapire
and the organizers and attendees of the CAGI 2007 Workshop on Grammar
Induction.

This material is based upon work 
supported by DARPA under grant FA8750-07-C-0014
and the NSF
   under grants 0612147 and 0615062.
Any opinions, findings, and conclusions or recommendations
   expressed in this material are those of the authors and do not
   necessarily reflect the views of DARPA or the NSF.


\bibliographystyle{plainnat}
\begin{thebibliography}{37}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Angluin(1982)]{angluin:revesible-language-inference}
Dana Angluin.
\newblock Inference of reversible languages.
\newblock \emph{Journal of the ACM}, 29\penalty0 (3):\penalty0 741--765, 1982.

\bibitem[Arasu and Garcia-Molina(2003)]{arasu+:sigmod03}
Arvind Arasu and Hector Garcia-Molina.
\newblock Extracting structured data from web pages.
\newblock In \emph{{SIGMOD}}, pages 337--348, 2003.

\bibitem[Bex et~al.(2006)Bex, Neven, Schwentick, and Tuyls]{bex+:dtd-inference}
Geert~Jan Bex, Frank Neven, Thomas Schwentick, and Karl Tuyls.
\newblock Inference of concise {DTDs} from {XML} data.
\newblock In \emph{{VLDB}}, pages 115--126, 2006.

\bibitem[Bex et~al.(2007)Bex, Neven, and
  Vansummeren]{bex+:inferring-xml-schema}
Geert~Jan Bex, Frank Neven, and Stijn Vansummeren.
\newblock Inferring {XML} schema definitions from {XML} data.
\newblock In \emph{{VLDB}}, pages 998--1009, 2007.

\bibitem[Borkar et~al.(2001)Borkar, Deshmukh, and
  Sarawagi]{borkar+:text-segmentation}
Vinayak Borkar, Kaustubh Deshmukh, and Sunita Sarawagi.
\newblock Automatic segmentation of text into structured records.
\newblock In \emph{{SIGMOD}}, pages 175--186, New York, NY, USA, 2001.

\bibitem[Burke et~al.(2007)Burke, Fisher, Walker, White, and
  Zhu]{burke+:cagi07}
David Burke, Kathleen Fisher, David Walker, Peter White, and Kenny~Q. Zhu.
\newblock Towards 1-click tool generation with {PADS}.
\newblock In \emph{{CAGI}}, Corvallis, OR, June 2007.

\bibitem[Chawathe et~al.(1994)Chawathe, Garcia-Molina, Hammer, Ireland,
  Papakonstantinou, Ullman, and Widom]{chawathe+:tsimmis}
Sudarshan Chawathe, Hector Garcia-Molina, Joachim Hammer, Kelly Ireland, Yannis
  Papakonstantinou, Jeffrey~D. Ullman, and Jennifer Widom.
\newblock The {TSIMMIS} project: Integration of heterogeneous information
  sources.
\newblock In \emph{16th Meeting of the Information Processing Society of
  Japan}, pages 7--18, Tokyo, Japan, 1994.

\bibitem[Crescenzi et~al.(2001)Crescenzi, Mecca, and
  Merialdo]{Cresenzi+:roadrunner}
Valter Crescenzi, Giansalvatore Mecca, and Paolo Merialdo.
\newblock Roadrunner: Towards automatic data extraction from large web sites.
\newblock In \emph{{VLDB}}, pages 109--118, San Francisco, CA, USA, 2001.

\bibitem[Denis et~al.(2004)Denis, Lemay, and
  Terlutte]{denis:learning-regular-languages}
Fran\c{c}ois Denis, Aur{\'e}lien Lemay, and Alain Terlutte.
\newblock Learning regular languages using {RFSAs}.
\newblock \emph{Theoretical Computer Science}, 313\penalty0 (2):\penalty0
  267--294, 2004.

\bibitem[Fern\'andez et~al.(2006)Fern\'andez, Fisher, Gruber, and
  Mandelbaum.]{fernandez+:padx}
Mary~F. Fern\'andez, Kathleen Fisher, Robert Gruber, and Yitzhak Mandelbaum.
\newblock {PADX}: Querying large-scale ad hoc data with {XQuery}.
\newblock In \emph{{PLAN-X}}, January 2006.

\bibitem[Fernau(2001)]{fernau:learning-xml}
Henning Fernau.
\newblock Learning {XML} grammars.
\newblock In \emph{{MLDM}}, pages 73--87, 2001.

\bibitem[Fisher and Gruber(2005)]{fisher+:pads}
Kathleen Fisher and Robert Gruber.
\newblock {PADS}: A domain specific language for processing ad hoc data.
\newblock In \emph{{PLDI}}, pages 295--304, June 2005.

\bibitem[Fisher et~al.(2006)Fisher, Mandelbaum, and Walker]{fisher+:popl06}
Kathleen Fisher, Yitzhak Mandelbaum, and David Walker.
\newblock The next 700 data description languages.
\newblock In \emph{{POPL}}, January 2006.

\bibitem[Garofalakis et~al.(2000)Garofalakis, Gionis, Rastogi, Seshadri, and
  Shim]{garofalakis+:xtract}
Minos~N. Garofalakis, Aristides Gionis, Rajeev Rastogi, S.~Seshadri, and
  Kyuseok Shim.
\newblock {XTRACT}: A system for extracting document type descriptors from
  {XML} documents.
\newblock In \emph{{SIGMOD}}, pages 165--176, 2000.

\bibitem[Gold(1967)]{gold:inference}
E.~M. Gold.
\newblock Language identification in the limit.
\newblock \emph{Information and Control}, 10\penalty0 (5):\penalty0 447--474,
  1967.

\bibitem[Gr\"{u}nwald(2007)]{mdlbook}
Peter~D. Gr\"{u}nwald.
\newblock \emph{The Minimum Description Length Principle}.
\newblock {MIT Press}, May 2007.

\bibitem[Hong(2002)]{hong:thesis}
Theodore~W. Hong.
\newblock \emph{Grammatical Inference for Information Extraction and
  Visualisation on the Web}.
\newblock Ph.D. Thesis, Imperial College London, 2002.

\bibitem[Huhtala et~al.(1999)Huhtala, K{\"a}rkk{\"a}inen, Porkka, and
  Toivonen]{TANE-HKPT99}
Yk{\"a} Huhtala, Juha K{\"a}rkk{\"a}inen, Pasi Porkka, and Hannu Toivonen.
\newblock {TANE}: An efficient algorithm for discovering functional and
  approximate dependencies.
\newblock \emph{The Computer Journal}, 42\penalty0 (2):\penalty0 100--111,
  1999.

\bibitem[Hutchens and Alder(1998)]{hutchens98finding}
Jason~L. Hutchens and Michael~D. Alder.
\newblock Finding structure via compression.
\newblock In David M.~W. Powers, editor, \emph{Proceedings of the Joint
  Conference on New Methods in Language Processing and Computational Natural
  Language Learning}, pages 79--82. 1998.

\bibitem[Kushmerick(1997)]{kushmerick-phd1997}
N.~Kushmerick.
\newblock \emph{Wrapper induction for information extraction}.
\newblock PhD thesis, University of Washington, 1997.
\newblock Department of Computer Science and Engineering.

\bibitem[Kushmerick et~al.(1997)Kushmerick, Weld, and
  Doorenbos]{KushmerickWD97:Wrapper}
Nicholas Kushmerick, Daniel~S. Weld, and Robert~B. Doorenbos.
\newblock Wrapper induction for information extraction.
\newblock In \emph{{IJCAI}}, pages 729--737, 1997.

\bibitem[Lerman et~al.(2004)Lerman, Getoor, Minton, and
  Knoblock]{Lerman+:webtables}
Kristina Lerman, Lise Getoor, Steven Minton, and Craig Knoblock.
\newblock Using the structure of web sites for automatic segmentation of
  tables.
\newblock In \emph{{SIGMOD}}, pages 119--130, New York, NY, USA, 2004.

\bibitem[Lin(1991)]{Lin91:divergence}
J.~Lin.
\newblock Divergence measures based on the {Shannon} entropy.
\newblock \emph{IEEE Transactions on Information Theory}, 37\penalty0
  (1):\penalty0 145--151, 1991.

\bibitem[Mandelbaum et~al.(2007)Mandelbaum, Fisher, Walker, Fernandez, and
  Gleyzer]{mandelbaum+:pads-ml}
Yitzhak Mandelbaum, Kathleen Fisher, David Walker, Mary Fernandez, and Artem
  Gleyzer.
\newblock {PADS/ML}: {A} functional data description language.
\newblock In \emph{{POPL}}, January 2007.

\bibitem[Martens et~al.(2006)Martens, Neven, Schwentick, and
  Bex]{martens+:expressiveness-xml-schema}
Wim Martens, Frank Neven, Thomas Schwentick, and Geert~Jan Bex.
\newblock Expressiveness and complexity of {XML} schema.
\newblock \emph{ACM Transactions on Database Systems}, 31\penalty0
  (3):\penalty0 770--813, 2006.

\bibitem[Muslea et~al.(2003)Muslea, Minton, and
  Knoblock]{muslea+:active-learning}
Ion Muslea, Steve Minton, and Craig Knoblock.
\newblock Active learning with strong and weak views: a case study on wrapper
  induction.
\newblock In \emph{{IJCAI}}, pages 415--420, 2003.

\bibitem[Ng et~al.(1999)Ng, Lim, and Koo]{Ng+:texttables}
Hwee~Tou Ng, Chung~Yong Lim, and Jessica Li~Teng Koo.
\newblock Learning to recognize tables in free text.
\newblock In \emph{{ACL}}, pages 443--450, Morristown, NJ, USA, 1999.

\bibitem[Oncina and Garcia(1992)]{rpni}
J.~Oncina and P.~Garcia.
\newblock Inferring regular languages in polynomial updated time.
\newblock \emph{Machine Perception and Artificial Intelligence}, 1:\penalty0
  29--61, 1992.

\bibitem[PADS Project()]{padsweb}
PADS Project.
\newblock {PADS} project.
\newblock {\url{http://www.padsproj.org/}}, 2007.

\bibitem[Pinto et~al.(2003)Pinto, McCallum, Wei, and Croft]{Pinto+:texttables}
David Pinto, Andrew McCallum, Xing Wei, and W.~Bruce Croft.
\newblock Table extraction using conditional random fields.
\newblock In \emph{SIGIR}, pages 235--242, New York, NY, USA, 2003.

\bibitem[Raeymaekers et~al.(2005)Raeymaekers, Bruynooghe, and den
  Bussche]{raeymaekers+:learning-tree-languages}
Stefan Raeymaekers, Maurice Bruynooghe, and Jan~Van den Bussche.
\newblock Learning (k, l)-contextual tree languages for information extraction.
\newblock In \emph{{ECML}}, pages 305--316, 2005.

\bibitem[Raman and Hellerstein(2001)]{raman+:potterwheel}
Vijayshankar Raman and Joseph~M. Hellerstein.
\newblock Potter's wheel: An interactive data cleaning system.
\newblock In \emph{{VLDB}}, pages 381 -- 390, 2001.

\bibitem[Shoens et~al.(1993)Shoens, Luniewski, Schwarz, Stamos, and
  Joachim~Thomas]{shoens+:rufus}
Kurt~A. Shoens, Allen Luniewski, Peter~M. Schwarz, James~W. Stamos, and
  II~Joachim~Thomas.
\newblock The {Rufus} system: Information organization for semi-structured
  data.
\newblock In \emph{{VLDB}}, pages 97--107, San Francisco, CA, USA, 1993.

\bibitem[Soderland(1999)]{soderland:whisk}
Stephen Soderland.
\newblock Learning information extraction rules for semi-structured and free
  text.
\newblock \emph{Machine Learning}, 34\penalty0 (1-3):\penalty0 233--272, 1999.

\bibitem[Stolcke and Omohundro(1994)]{stolcke94inducing}
Andreas Stolcke and Stephen Omohundro.
\newblock Inducing probabilistic grammars by bayesian model merging.
\newblock In \emph{{ICGI}}, pages 106--118, 1994.

\bibitem[Vidal(1994)]{vidal:gisurvey}
Enrique Vidal.
\newblock Grammatical inference: An introduction survey.
\newblock In \emph{{ICGI}}, pages 1--4, 1994.

\bibitem[Wharton(1974)]{wharton:approximate-language-identification}
R.~M. Wharton.
\newblock Approximate language identification.
\newblock \emph{Information and Control}, 26\penalty0 (3):\penalty0 236--255,
  1974.

\end{thebibliography}
\end{document}


