\documentclass{entcs} 
\usepackage{entcsmacro}
\usepackage{graphicx}
\usepackage{code,amsmath,amssymb}
\usepackage{math-cmds}

\include{defs}

% A couple of exemplary definitions:
\def\lastname{Fernandez, Fisher, Mandelbaum and Walker}
\begin{document}
\begin{frontmatter}
  \title{A Language for Describing and Transforming Ad Hoc Data} 
  \author{Mary Fern\'andez,
    Kathleen Fisher\thanksref{attemail}}
  \address{AT\&T\\ 
    Florham Park,NJ USA} 
  \author{Yitzhak Mandelbaum,
    David Walker\thanksref{premail}}
  \address{Department of Computer Science\\ 
    Princeton University\\
    Princeton,NJ USA} 
  \thanks[attemail]{Email:\texttt{\normalshape
        mff,kfisher@research.att.com}}
  \thanks[premail]{Email:\texttt{\normalshape
        yitzhakm,dpw {\it at} cs {\it dot} princeton {\it dot} edu}}
\begin{abstract}
\datatype{} pairs a declarative data description language with an 
error-aware transformation language to provide rich support for
programming with ad hoc data sources.  Such data is common in a wide range of domains including networking, financial analysis, biology, and physics.
\datatype{}'s data description language is based on polymorphic, recursive, and dependent data types.  From such descriptions, the compiler generates robust parsing and pretty printing code. The transformation language supports \textit{error-aware computation} by automatically maintaining an association between parsed data and a description of the errors in that data.  This association supports an obliviousness property, which allows programmers to specify transforms for correct data, trusting the system to preserve error tags during such transformations.
This reification of errors allows programmers to filter, query, and repair errors on the one hand, while allowing programs to safely ignore errors on the other.  

We present the design of an \ml{}-like language with intrinsic support for parsing, transforming, and pretty printing \textit{ad hoc} data. We use the term  ad hoc data to refer to any non-standard data source.  Such data is common 

  
   language for describing and transforming ad hoc data,
  that is, data in a non-standard format. The data descriptions are in
  the form of types based closely on \ml{}-style types, including
  parametric datatypes, type definitions, and records. They are used
  by the compiler to generate parsing functions from the data source
  to special internal language data structures, called \pvalue{}s.
%   Furthermore, these data structures are combinations of
%   parsed data and parse descriptors, which describe properties of the
%   parsing process, including errors. 
  The language provides new patterns and data constructors for 
  transforming \pvalue{}s.
\end{abstract}
\begin{keyword}
  Ad hoc data, data description, data transformation, pattern matching, 
  dependent types.
\end{keyword}
\end{frontmatter}

\input{intro}

\input{descr}

\input{trans}

\input{implementation}

\input{relatedwork}

\section{Conclusions and Future Work}
\label{sec:conclusion}

\bibliographystyle{abbrv}
\bibliography{pads}
\end{document}
