\section{Related Work}
\label{sec:related}

The work in this paper builds upon ideas developed in the \pads{} 
project~\cite{fisher+:pads,fisher+:toplas}. In \pads{}, extended
type declarations are used to describe the grammar of a document
and simultaneously to generate types for parsed data as well as a suite 
of data-processing tools.  The obvious difference between 
\pads{} (or other parser generators) and
\forest{} is that 
\pads{} generates infrastructure for processing strings (the insides
of a single file) whereas \forest{} generates infrastructure for 
processing entire file systems.
In addition,
architecturally, \forest{} (and \padshaskell) is superior to 
previous versions of \pads{}
in its support for third-party generic programming and tool construction.

More generally, \forest{} shares high-level goals with other systems
that seek to make data-oriented programming simpler and more productive.
For example, Microsoft's LINQ~\cite{linq} extends the .NET languages
by making it possible to query 
any datasource that supports the IEnumerable interface using
a simple, convenient syntax.  However, LINQ
differs from \forest{} in that
it does not provide declarative 
language-integrated syntax for specifying the structure of, and then ingesting,
external ad hoc \filestores{}.

{\em Type Providers}~\cite{syme+:type-providers}, an experimental feature
of F\#, help programmers materialize standard data sources equipped with
predefined schema (such as \xml{} documents or databases) in memory in
an F\# program.  Type Providers and \forest{} descriptions
are complementary language
features.  In fact, it may be possible to define a new F\# Type Provider
capable of interpreting \forest{} file system schema and ingesting
the described data, thereby making \forest{}-described data available
in F\#.

In the databases community, a number of XML-based
description languages have been defined for specifying
file format, file organization and file location.
One example of such a language is XFiles~\cite{xml-file-sys}.
XFiles has many features in common with \forest{}:  It can
describe file locations, permissions, ownership and other
attributes.  It can also specify the name of an application capable of
parsing the files in question.  The main difference between
a language like XFiles and \forest{} is that \forest{} is
tightly integrated into a general-purpose, conventional programming
language.  \forest{} declarations generate types, functions
and data structures that materialize within
a greater \haskell{} program; XFiles 
does not interoperate directly with conventional programming languages.
