\subsection{Future Naval Relevance}
\label{sec:naval}

Naval information systems involve a complex, heterogenous
infrastructure with world-wide reach and a vast ecosystem of
interacting, distributed applications. Going forward, as more
information is collected and disseminated to more participants, the
distributed nature of applications will become even more ingrained in
these systems. Given the inherent complexity in distributed systems,
and the difficulty of diagnosing problems in them, monitoring and
proactive detection are necessary to keep this ecosystem functioning
smoothly.  However, as systems and applications diversify, building
appropriate defensive monitoring infrastructure becomes more
difficult, time-consuming, error-prone, and, ultimately, more costly.

We believe that the best hope for building trustworthy, high
reliability systems in such complex future environments is to develop
new technology capable of partially or fully automating the
construction of monitoring software and systems.  Our research will
help proactively find problems, record/archive system status, oversee
operations, debug new applications and detect malicious processes.  In
order to do this, we have proposed a multi-pronged approach that
combines ideas from the programming languages and systems community:

\begin{itemize}
\item Prong 1:  Develop a new programming language capable of specifying the 
form and properties of data produced by networked systems and 
applications.  {\em Automatically} compile those specifications into 
components for parsing, traversal, transformation, error detection,
statistical analysis and querying needed by monitoring systems.

\item Prong 2:  Provide {\em automatic} support for generating the data 
specifications by inferring them directly from available data.

\item Prong 3:  Build new systems support for {\em automatically}
isolating live traffic flows, associating them with their applications,
learning normal behavior, and detecting anomalous traffic.

\item Prong 4:  Wrap all previous components up in a general, scalable,
and adaptive monitoring system capable of
presenting new and evolving data sources to operators and allowing them to
analyze and visualize data across multiple domains.
\end{itemize}

Together, these four prongs form a blueprint for development of the
next generation of monitoring infrastructure.