\newif\ifdraft\drafttrue
\newif\ifcolor\colortrue

% For per-person control of tex'ing, put commands like \twocolfalse
% in a file called texdirectives.tex, which we read at this point (if
% it exists).  Note that this file should be left out of the SVN
% repository. 
\makeatletter \@input{texdirectives} \makeatother

\documentclass[nocopyrightspace]{sigplanconf}

\usepackage{alltt}
\usepackage{balance}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{code}
\usepackage{color}
\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{url}
\usepackage{verbatim}
\usepackage{pervasives}

\newcommand{\cut}[1]{}
\newcommand{\note}[1]{\textbf{Note:}#1}

\newcommand{\appref}[1]{Appendix~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\tblref}[1]{Table~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\listingref}[1]{Listing~\ref{#1}}
\newcommand{\bftt}[1]{{\ttfamily\bfseries{}#1}}
\newcommand{\kw}[1]{\bftt{#1}}
\newcommand{\eg}{{\em e.g.}}
\newcommand{\cf}{{\em cf.}}
\newcommand{\ie}{{\em i.e.}}
\newcommand{\etc}{{\em etc.\/}}

\newcommand{\hscomment}[1]{{\color{red}\{- #1 -\}}}


\title{Transactional Forest}

\authorinfo{Submission \#274}{}{\vspace*{-4cm}}

\begin{document}

\maketitle

\begin{abstract}
Many applications rely on the file system to store persistent data,
but current programming languages lack convenient constructs for
manipulating file system data. Previous work on the Forest language
developed a type-based abstraction for file systems in which the
programmer writes a high-level specification describing the expected
structure of the file system, and the compiler generates an in-memory
representation for the data and accompanying ``load'' and ``store''
functions. Unfortunately Forest does not provide any consistency
guarantees so if multiple applications are manipulating the file
system concurrently---by far the common case---it can produce
incorrect results.

This paper presents Transactional Forest: an extension to Forest that
enriches the language with seralizable transactions. We present the
design of the language, which is based on a new ``atomic'' construct
and a monad that tracks effects. We formalize the semantics of POSIX
file systems in a simple core calculus and prove the correctness of
our implementation. We discuss our implementation in Haskell and
illustrate its use on a substantial case study: the Soil and Water
Assessment Tool (SWAT), which is a modeling tool used by numerous
hydrologists and environmental scientists.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

\input{introduction}

\begin{itemize}
\item Background
\begin{itemize}
\item Forest Language (primitives, load, store)
\item Running example 
\item Problems with naive semantics
\item Transactions to the rescue
\end{itemize}
\item TxForest Language
\begin{itemize}
\item Atomic construct 
\item Transaction monad
\item Varieties of failure
\item Revised running example
\item Guarantees
\end{itemize}
\item Featherweight POSIX
\begin{itemize}
\item Discussion of Sewell-eque formalism vs. core calculi  
\item Showcase subtlety (e.g., weak locking primitives?)
\item Define translation from TxForest to IMPOSIX (defines TxForest's semantics)
\item Reference implementations (data structure locks, lockf, etc.)
\item Prove serializability for fully Forested programs
\end{itemize}
\item Implementation
\item SWAT Case Study
\item Related Work
\item Conclusion
\end{itemize}

\input{background}

\input{txforest}

\section{Featherweight Posix}
\label{sec:posix}

\input{posix}
\input{implementation}
\input{case-study}
\input{related}
\input{conclusion} 
\bibliographystyle{plain} 
\balance  
\bibliography{main}

\section*{Appendix}
\label{sec:appendix}
\input{appendix}

\end{document}
