%\documentclass[fleqn]{article}
\documentclass[nocopyrightspace]{sigplanconf}

\usepackage{xspace,pads,amsmath,math-cmds,
            math-envs,inference-rules,times,
            verbatim,alltt,multicol,proof,url}

\usepackage{code} 
\usepackage{epsfig}

%\setlength{\oddsidemargin}{0in}
%\setlength{\evensidemargin}{0in}
%\setlength{\textwidth}{6.5in}
%\setlength{\textheight}{8.5in}

\begin{document}
\title{PADS/ML: A Functional Data Description Language}
\authorinfo{Yitzhak Mandelbaum$^\ast$, Kathleen Fisher$^\dagger$, David Walker$^\ast$, Mary
  Fernandez$^\dagger$, Artem Gleyzer$^\ast$}
{$^\ast$Princeton University \ \ \ \ \ \ \ $^\dagger$AT\&T Labs Research}
 {\mono{yitzhakm,dpw,agleyzer@CS.Princeton.EDU} \ \ \ \ \mono{kfisher,mff@research.att.com}}

\input{definitions}

\maketitle{}

\begin{abstract}  

  Massive amounts of useful data are stored and processed in
  \textit{ad hoc} formats for which common tools like parsers and
  pretty printers do not exist.  Traditional data management systems
  provide rich infrastructure for processing well-behaved data, but
  are of little use when dealing with data in ad hoc formats.  To
  address the challenges of ad hoc data, we have designed \padsml{}, a
  declarative data description language for the \ml{} family of
  languages.  \padsml{} is based on the \ml{} type structure and
  features polymorphic, dependent, recursive datatypes for describing
  ad hoc data.  In addition, \padsml{} uses the \ml{} module system as
  a general framework to help users easily develop customized
  data-processing tools for any data format described in \padsml{}.
  We have formalized the semantics of \padsml{}'s datatypes by
  extending our previous work on the Data Description
  Calculus~\cite{fisher+:next700ddl} with type-parameterized types,
  and we have proven the resulting system ``type-correct,'' \ie{},
  generated parsers return data of the expected type.  Finally, we
  have implemented a version of \padsml{} for the \ocaml{} language.
  From a \padsml{} description, our compiler generates an \ocaml{}
  module containing the relevant type declarations, a parser for the
  description, and other tools.

\end{abstract}

\input{intro}

\input{pml_overview}

\input{impl}

\input{gentool}

\input{ddcpoly}

\input{related_work}

\input{conc}

\section*{Acknowledgments}

We would like to thank Derek Dreyer for discussions and advice on
advanced module systems.

\bibliographystyle{abbrv}
\bibliography{pads}
%\bibliography{../common/pads-long,../common/pads}

\input{appendix}

\end{document}

%%% Local Variables:
%%% mode: outline-minor
%%% End:

